### 2015-02-28 writing/F'14 paper/intro.tex
Paper update

The experiments with simulated data provide a convincing demonstration
of the ability of neural networks to learn to encode semantic
inferences in complex natural language sentences from reasonably-sized
training sets. However, we are also interested in the more practical
question of whether they can learn these representations from
naturalistic text. To address this question, we apply our models to
the SICK entailment challenge data in \S\ref{sec:sick}. The small size of this corpus puts
data-hungry NN models like ours at a disadvantage, but we are
nonetheless able to achieve competitive performance on it,
surpassing several submitted models with significant hand-engineered task-specific features.
Our concern is primarily with the quality of the learned representations and their ability to encode our target semantic relations.
Thus, we are less focused on absolute performance than on patterns of accurate and inacccurate predictions, so even this small data set can offer insight into our core representational issues.
This suggests that the representational abilities that we observe in the previous sections
are not limited to carefully circumscribed tasks. We conclude that TreeRNTN models
are adequate for typical cases of natural logic inference, and that there is not yet
any clear level of inferential complexity for which other approaches work and NN models fail.

>>>>

The experiments with simulated data provide a convincing demonstration
of the ability of neural networks to learn to build and use semantic representations for complex natural language sentences from reasonably-sized
training sets. However, we are also interested in the more practical
question of whether they can learn these representations from
naturalistic text. To address this question, we apply our models to
the SICK entailment challenge data in \S\ref{sec:sick}. The small size of this corpus puts
data-hungry NN models like ours at a disadvantage, but we are
nonetheless able to achieve competitive performance on it,
surpassing several submitted models with significant hand-engineered task-specific features and our own NN baseline.
This suggests that the representational abilities that we observe in the previous sections
are not limited to carefully circumscribed tasks. We conclude that TreeRNTN models
are adequate for typical cases of natural language inference, and that there is not yet
any clear level of inferential complexity for which other approaches work and NN models fail.

### 2015-02-28 writing/F'14 paper/methods.tex
Paper update.
Paper update

\section{Recursive neural network models} \label{methods}

We limit the scope of our experiments in this paper to neural network models 
that adhere to the linguistic \ii{principle of
 compositionality}, which says that the meanings for complex
expressions are derived from the meanings of their parts
via specific composition functions \cite{Partee84,Janssen97}. In our
distributed setting, word meanings are embedding vectors of dimension $n$. The
composition function maps pairs of them to single phrase vectors of dimension $n$, 
which can then be merged again to represent more complex
phrases. Once the entire sentence-level representation has been
derived, it serves as a fixed-dimensional input for some subsequent function.

@@

Other strong tree-structured networks have been proposed in past work
\cite{sochergrounded,kalchbrenner2014convolutional,irsoydeep}, but
we believe that these two provide a valuable case study, and that positive results on 
here are likely to generalize well to stronger models.

To run the model forward, we assemble the two tree networks so as to match the structures provided for each phrase, which is either included in the source data or given by a parser.
The word vectors are then looked up from the vocabulary embedding matrix $V$ (one of the model parameters), and
the composition and comparison functions are used to pass information
up the tree and into the classifier. For an objective
function, we use the negative log of the probability assigned to the
correct label with L2 regularization.

We initialize parameters uniformly, using the range $(-0.05, 0.05)$ for layer parameters
and $(-0.01, 0.01)$ for embeddings, and train the model using stochastic gradient descent (SGD)
with learning rates computed using AdaDelta \cite{zeiler2012adadelta}.
The classifier feature vector is fixed at 75 dimensions and 
the dimensions of the recursive layers are tuned manually.
Training times on CPUs vary from hours to days across experiments.
On the first three experiments, which use artificial data, we report mean
results over five fold cross-validation.
Variance across runs is typically no
more than two percentage points. In addition, because the classes are not necessarily balanced, 
we report performance 
using both accuracy and macroaveraged F1.\footnote{We compute macroaveraged F1 
as the harmonic mean of average precision and average recall, both computed
for all classes for which there is test data, setting precision to 0 
where it is not defined.}
Source code and generated data will be released after the review period.

>>>>

\section{Tree-structured neural networks} \label{methods}

We limit the scope of our experiments in this paper to neural network models 
that adhere to the linguistic \ii{principle of
 compositionality}, which says that the meanings for complex
expressions are derived from the meanings of their parts
via specific composition functions \cite{Partee84,Janssen97}. In our
distributed setting, word meanings are embedding vectors of dimension $n$. A learned
composition function maps pairs of them to single phrase vectors of dimension $n$, 
which can then be merged again to represent more complex
phrases, forming a tree structure. Once the entire sentence-level representation has been
derived at the top of the tree, it serves as a fixed-dimensional input for some subsequent layer function.

@@

Other strong tree-structured models have been proposed in past work
\cite{sochergrounded,kalchbrenner2014convolutional,irsoydeep}, but
we believe that these two provide a valuable case study, and that positive results on 
here are likely to generalize well to stronger models.

To run the model forward, we assemble the two tree-structured networks so as to match the structures provided for each phrase, which are either included in the source data or given by a parser.
The word vectors are then looked up from the vocabulary embedding matrix $V$ (one of the learned model parameters), and
the composition and comparison functions are used to pass information
up the tree and into the classifier. For an objective
function, we use the negative log of the probability assigned to the
correct label with L2 regularization.

We initialize parameters uniformly, using the range $(-0.05, 0.05)$ for layer parameters
and $(-0.01, 0.01)$ for embeddings, and train the model using stochastic gradient descent (SGD)
with learning rates computed using AdaDelta \cite{zeiler2012adadelta}.
The classifier feature vector is fixed at 75 dimensions and 
the dimensions of the recursive layers are tuned manually.
Training times on CPUs vary from hours to days across experiments.
On the experiments which use artificial data, we report mean
results over five fold cross-validation, where
variance across runs is typically no
more than two percentage points. In addition, because the classes are not necessarily balanced, 
we report both accuracy and macroaveraged F1.\footnote{We compute macroaveraged F1 
as the harmonic mean of average precision and average recall, both computed
for all classes for which there is test data, setting precision to 0 
where it is not defined.}
Source code and generated data will be released after the review period.

### 2015-02-28 writing/F'14 paper/quantifiers.tex
Paper update

We have seen that recursive models can learn an approximation of propositional
logic.  However, natural languages can express functional meanings of
considerably greater complexity than this.  As a key test of whether our models 
can capture this complexity, we now directly measure the degree to which they are able to
develop suitable representations for the semantics of natural language
quantifiers like \ii{most} and \ii{all} as they interact with negation. Quantification 
and negation are far from the only place in natural language where complex functional meanings
are found, but they are natural focus, since they have
formed a standard case study in prior formal work on natural
language inference \cite{Icard:Moss:2013:LILT}.

@@

\vspace{-0.6cm}
\begin{gather}
  \text{(most turtle) swim} \natalt \text{(no turtle) move}\label{p1}
  \\
  \text{(all lizard) reptile} \natfor  \text{(some lizard) animal}\label{p2}
  \\
  \text{(most turtle) reptile} \natalt \text{(no turtle) animal}\label{p3}
\end{gather}

We use the same summing baseline as in \S\ref{sec:recursion}.
The highly consistent  sentence structure in this experiment means
that that model can potentially distinguish two valid sentences unless they
 differ only in the placement of \ii{not}, potentially introducing errors on 
the 58.8\% of examples that contain a sentence with a single
instance of \ii{not}.

>>>>

We have seen that recursive models can learn an approximation of propositional
logic.  However, natural languages can express functional meanings of
considerably greater complexity than this.  As a key test of whether our models 
can capture this complexity, we now study the degree to which they are able to
develop suitable representations for the semantics of natural language
quantifiers like \ii{most} and \ii{all} as they interact with negation and lexical entailments. Quantification 
and negation are far from the only place in natural language where complex functional meanings
are found, but they are natural focus, since they have
formed a standard case study in prior formal work on natural
language inference \cite{Icard:Moss:2013:LILT}.

@@

\vspace{-0.6cm}
\begin{gather}
  \text{(most turtle) swim} \natalt \text{(no turtle) move}\label{p1}
  \\
  \text{(all lizard) reptile} \natfor  \text{(some lizard) animal}\label{p2}
  \\
  \text{(most turtle) reptile} \natalt \text{(all turtle) (not animal)}\label{p3}
\end{gather}

We use the same summing baseline as in \S\ref{sec:recursion}.
The highly consistent  sentence structure in this experiment means that this model
is not as disadvantaged by the lack of word order information as it is in the previous experiment, 
but the variable placement of \ii{not}  nonetheless introduces potential uncertainty in the 58.8\% 
of examples that contain a sentence with a single token of it.

### 2015-02-28 writing/F'14 paper/sick.tex
Paper update

The specific model architecture that we use is novel, and though the underlying tree structure approach has been validated elsewhere, our experiments so far do not guarantee that it viable model for handling inference over real
natural language data. To investigate our models' ability to handle the noisy labels and the diverse range of linguistic structures seen in typical natural language data, we train a version of our model on the SICK textual entailment challenge corpus \cite{marelli2014sick}. The corpus consists of about 10k natural language sentence pairs, labeled with \ii{entailment}, \ii{contradiction}, or \ii{neutral}. At only a few thousand distinct sentences (many of them variants on an even smaller set of template sentences), the corpus is not large enough to train a high quality learned model of general natural language, but it is the largest human-labeled entailment corpus that we are aware of, and our results
nonetheless show that tree-structured NN models can learn to do inference in the real world.

Adapting to this task required us to make a few additions to the techniques discussed in \S\ref{methods}. In order to better handle rare words, we initialized our word embeddings using 200 dimensional vectors trained with 
GloVe \cite{pennington2014glove} on data from Wikipedia. Since 200 dimensional vectors are too large to be practical in an TreeRNTN on a small dataset, a new embedding transformation layer is needed. Before any embedding is used as an input to a recursive layer, it is passed through an additional $\tanh$ neural network layer with the same output dimension as the recursive layer. This new layer aggregates any usable information from the embedding vectors into a more compact working representation. An identical layer is added to the SumNN between the word vectors and the comparison layer.

We also supplemented the SICK training data\footnote{We tuned the model using performance on a held out development set, but report performance here for a version of the model trained on both the training and development data and tested on the SICK test set. We also report training accuracy on a small sample from each data source.} with 600k examples of entailment data from the Denotation Graph project (DG, \citealt{hodoshimage}, also used by the winning SICK submission), a corpus of noisy automatically labeled entailment examples over image captions, the same genre of text from which SICK was drawn. We trained a single model on data from both sources, but used a separate set of softmax parameters for classifying into the labels from each source. We parsed the data from both sources with the Stanford PCFG Parser v.~3.3.1 \cite{klein2003accurate}. We also found that we were able to train a working model much more quickly with an additional technique: we collapse subtrees that were identical across both sentences in a pair by replacing them with a single head word. The training and test data on which we report performance are collapsed in this way, and both collapsed and uncollapsed copies of the training data are used in training. Finally, in order to improve regularization on the noisier data, we used dropout \cite{srivastava2014dropout} at the input to the comparison layer (10\%) and at the output from the embedding transform layer (25\%).

\begin{table}[tp]
  \centering \small
    \begin{tabular}{ l@{~~} r@{~~~~} r@{~~~~} r@{~~~~} r@{~~~~} }
    \toprule
        ~&\ii{neut.}&	 30d  & 			30d & 50d\\
    ~&only &SumNN  &TrRNN &TrRNTN\\ 
     \midrule
    DG Train	& 50.0 & 64.0 & 67.0 & \textbf{74.0} \\
    SICK Train	& 56.7 & 95.2 & 95.4 & \textbf{97.8} \\
    SICK Test	& 56.7 & 73.6 & 74.9 & \textbf{76.9} \\
    \midrule
    \textsc{Passive} (4\%)	& 0 		& 80  		& 68		&\textbf{88}\\   
    \textsc{Neg} (7\%)	& 0 		& 88	 		& \textbf{100} & \textbf{100}\\
    \textsc{Subst} (24\%)	& 28 		& 60  		& 64 		&  \textbf{72}\\
    \textsc{MultiEd} (39\%)	&  \textbf{68} & 66  		&66 		& 64 \\
    \textsc{Diff} (26\%)	& \textbf{96} &  	82		&79		& \textbf{96}\\  
    \midrule
    \textsc{Short} (47\%) & 50.0 & 73.9 & 75.9		& \textbf{77.3} \\    
    \bottomrule
  \end{tabular}
  \caption{Classification accuracy, including a category breakdown for SICK test data. Categories are shown with their frequencies.}
  \label{sresultstable}
\end{table}

\begin{table*}[htp]
  \centering\small
  \begin{tabular}{lcl}
    \toprule
    A man is cutting a paper plate	& \ii{entailment} & A paper plate is being cut by the man (\textsc{Passive})\\
    A little girl is playing the violin on a beach & \ii{contradiction} &	There is no girl playing the violin on a beach (\textsc{Neg})\\
    A man is fixing a silencer to a gun & \ii{entailment} & A man is fitting a silencer to a weapon (\textsc{Subst})\\
        A woman is breaking two eggs in a bowl & \ii{neutral} &A man is mixing a few ingredients in a bowl (\textsc{MultiEd})\\
        Dough is being spread by a man & \ii{neutral} & A woman is slicing meat with a knife (\textsc{Diff})\\
    \bottomrule
  \end{tabular}
  \caption{\label{examplesofsickdata}Examples of each category used in error analysis from the SICK test data. }
\end{table*}

@@

We know that it is possible to perform well on SICK (like other natural language inference corpora) without taking advantage of compositional syntactic or semantic structure \cite{marelli2014semeval}, and our summing baseline model is powerful enough to do this. Our tree models nonetheless perform substantially better, and we remain confident that given sufficient data, it should be possible for the tree models, and not the summing model, to learn a truly high-quality solution.

To better understand how the models differ, we manually annotated a fraction of the SICK test set, using mutually exclusive categories for passive/active alternation pairs (\textsc{Passive}), pairs differing only by the presence of negation (\textsc{Neg}), pairs differing by a single word or phrase substitution (\textsc{Subst}), pairs differing by multiple edits (\textsc{Mult}), and pairs with little or no content word overlap (\textsc{Diff}). Examples of each are in Table \ref{examplesofsickdata}. We annotated 100 random examples to judge the frequency of each category, and  continued selectively annotating until each categogy contained at least 25. We also use category (\textsc{Short}) for pairs in which neither sentence contains more than ten words.

The results (Table \ref{examplesofsickdata}) show that the TreeRNTN performs especially strongly in the two categories which most narrowly pick out a specific syntactic configuration, \textsc{Passive} and \textsc{Neg}. It also beats the summing baseline on \textsc{Subst}, which most closely parallels the kind of monotonicity inferences studied in \S\ref{sec:quantifiers}. None of the three models perform better on the short \textsc{Short} pairs than on the rest of the data, suggesting that the decay phenomenon observed in \S\ref{sec:recursion} may not hurt the performance of models trained on typical natural language sentences.

>>>>

% TODO: Something on how we can pack enough

The specific model architecture that we use is novel, and though the underlying tree structure approach has been validated elsewhere, our experiments so far do not guarantee that it viable model for handling inference over real
natural language data. To investigate our models' ability to handle the noisy labels and the diverse range of linguistic structures seen in typical natural language data, we use the SICK textual entailment challenge corpus \cite{marelli2014sick}. The corpus consists of about 10k natural language sentence pairs, labeled with \ii{entailment}, \ii{contradiction}, or \ii{neutral}. At only a few thousand distinct sentences (many of them variants on an even smaller set of template sentences), the corpus is not large enough to train a high quality learned model of general natural language, but it is the largest human-labeled entailment corpus that we are aware of, and our results nonetheless show that tree-structured NN models can learn to do inference in the real world.

Adapting to this task requires us to make a few additions to the techniques discussed in \S\ref{methods}. In order to better handle rare words, we initialized our word embeddings using 200 dimensional vectors trained with 
GloVe \cite{pennington2014glove} on data from Wikipedia. Since 200 dimensional vectors are too large to be practical in an TreeRNTN on a small dataset, a new embedding transformation layer is needed. Before any embedding is used as an input to a recursive layer, it is passed through an additional $\tanh$ neural network layer with the same output dimension as the recursive layer. This new layer aggregates any usable information from the embedding vectors into a more compact working representation. An identical layer is added to the SumNN between the word vectors and the comparison layer.

We also supplemented the SICK training data\footnote{We tuned the model using performance on a held out development set, but report performance here for a version of the model trained on both the training and development data and tested on the 4,928 example SICK test set. We also report training accuracy on a small sample from each data source.} with 600k examples of entailment data from the Denotation Graph project (DG, \citealt{hodoshimage}, also used by the winning SICK submission), a corpus of noisy automatically labeled entailment examples over image captions, the same genre of text from which SICK was drawn. We trained a single model on data from both sources, but used a separate set of softmax parameters for classifying into the labels from each source. We parsed the data from both sources with the Stanford PCFG Parser v.~3.3.1 \cite{klein2003accurate}. We also found that we were able to train a working model much more quickly with an additional technique: we collapse subtrees that were identical across both sentences in a pair by replacing them with a single head word. The training and test data on which we report performance are collapsed in this way, and both collapsed and uncollapsed copies of the training data are used in training. Finally, in order to improve regularization on the noisier data, we used dropout \cite{srivastava2014dropout} at the input to the comparison layer (10\%) and at the output from the embedding transform layer (25\%).

\begin{table}[tp]
  \centering \small
    \begin{tabular}{ l@{\hspace{-0.25em}} r@{~~~~} r@{~~~~} r@{~~~~} r@{~~~~} }
    \toprule
        ~&\ii{neutral}&	 30d  & 			30d & 50d\\
    ~&only &SumNN  &TrRNN &TrRNTN\\ 
     \midrule
    DG Train	& 50.0 & 64.0 & 67.0 & \textbf{74.0} \\
    SICK Train	& 56.7 & 95.2 & 95.4 & \textbf{97.8} \\
    SICK Test	& 56.7 & 73.6 & 74.9 & \textbf{76.9} \\
    \midrule
    \textsc{Passive} (4\%)	& 0 		& 80  		& 68		&\textbf{88}\\   
    \textsc{Neg} (7\%)	& 0 		& 88	 		& \textbf{100} & \textbf{100}\\
    \textsc{Subst} (24\%)	& 28 		& 60  		& 64 		&  \textbf{72}\\
    \textsc{MultiEd} (39\%)	&  \textbf{68} & 66  		&66 		& 64 \\
    \textsc{Diff} (26\%)	& \textbf{96} &  	82		&79		& \textbf{96}\\  
    \midrule
    \textsc{Short} (47\%) & 50.0 & 73.9 & 75.9		& \textbf{77.3} \\    
    \bottomrule
  \end{tabular}
  \caption{Classification accuracy, including a category breakdown for SICK test data. Categories are shown with their frequencies.}
  \label{sresultstable}
\end{table}

\begin{table*}[htp]
  \centering\small
  \begin{tabular}{l@{~~~}cl}
    \toprule
  The patient is being helped by the doctor	& \ii{entailment} & The doctor is helping the patient (\textsc{Passive})\\
    A little girl is playing the violin on a beach & \ii{contradiction} &	There is no girl playing the violin on a beach (\textsc{Neg})\\

    The yellow dog is drinking water from a bottle& \ii{contradiction} &	The yellow dog is drinking water from a pot  (\textsc{Subst})\\
        A woman is breaking two eggs in a bowl & \ii{neutral} &A man is mixing a few ingredients in a bowl (\textsc{MultiEd})\\
        Dough is being spread by a man & \ii{neutral} & A woman is slicing meat with a knife (\textsc{Diff})\\
    \bottomrule
  \end{tabular}
  \caption{\label{examplesofsickdata}Examples of each category used in error analysis from the SICK test data. }
\end{table*}

@@

To better understand our results, we manually annotated a fraction of the SICK test set, using mutually exclusive categories for passive/active alternation pairs (\textsc{Passive}), pairs differing only by the presence of negation (\textsc{Neg}), pairs differing by a single word or phrase substitution (\textsc{Subst}), pairs differing by multiple edits (\textsc{MultiEd}), and pairs with little or no content word overlap (\textsc{Diff}). Examples of each are in Table \ref{examplesofsickdata}. We annotated 100 random examples to judge the frequency of each category, and  continued selectively annotating until each category contained at least 25. We also use the category \textsc{Short} for pairs in which neither sentence contains more than ten words.

The results (Table \ref{examplesofsickdata}) show that the TreeRNTN performs especially strongly in the two categories which pick out specific syntactic configurations, \textsc{Passive} and \textsc{Neg}, suggesting that that model has learned to encode the relevant structures well. It also beats both other models on \textsc{Subst}, which most closely parallels the lexical entailment inferences addressed in \S\ref{sec:quantifiers}. In addition none of the models perform better on the \textsc{Short} pairs than on the rest of the data, suggesting that the performance decay observed in \S\ref{sec:recursion} may not impact models trained on typical natural language text.

It is known that a model can perform well on SICK (like other natural language inference corpora) without taking advantage of compositional syntactic or semantic structure \cite{marelli2014semeval}, and our summing baseline model is powerful enough to do this. Our tree models nonetheless perform substantially better, and we remain confident that given sufficient data, it should be possible for the tree models, and not the summing model, to learn a truly high-quality solution.
